---
title: "Tim Book's R Utilities (`rutilstb`)"
author: "Timothy K. Book"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is a package of various functions that I've made which I've found useful
while working on different projects.  They fall into to two categories:

1. Functions I've found convenient while using Stata, and felt a need to port
them into R.
1. Functions I've needed for a specific, narrow purpose.

```{r, echo = FALSE}
library(rutilstb)
```

## Functions Ported from Stata
### The `isid` Function
The `isid` function can be used to detect whether or not a variable (or group of
variables) has unique values.  It has three uses:

#### Use #1: Vectors
```{r}
a <- 1:5
b <- c(1, 1, 2, 3, 4)
```
```{r}
isid(a)
isid(b)
```
#### Use #2: Data Frames
```{r}
users <- data.frame(
  user_id = 1:5,
  first_name = c('Tim', 'Brandon', 'Brandon', 'Arya', 'Cersei'),
  last_name = c('Book', 'Stark', 'Tully', 'Stark', 'Lannister')
)
```
```{r}
isid(users)
```
#### Use #3: Combinations of Variables in a Data Frame
```{r}
isid(users, keys = c('first_name', 'last_name'))
```

### The `tabstat` Function
The `tabstat` function can be used to view statistics based on categorical
variables within a data frame.
```{r}
tabstat('mpg', mtcars, by = 'cyl')
tabstat('mpg', mtcars, by = 'cyl', fns = list(median, IQR, length))
```

### The `mdesc` Function
The `mdesc` function gives a description of missing values in a data frame, and
where they are located.
```{r}
df <- data.frame(
  a = 1:5,
  b = c(2, 3, NA, 4, 5),
  c = c(NA, 'x', 'y', 'z', NA)
)
mdesc(df)
```

## Quick Functions of Narrow Use
### The `colMatcher` Function
The `colMatcher` function tells the user if any two columns are identical.
```{r}
df <- data.frame(
  x = 1:5,
  y = 101:105,
  z = 1:5,
  w = 101:105,
  v = letters[1:5]
)
colMatcher(df)
colMatcher(df, return_index = TRUE)
```

### The `dfSplit` Function
The `dfSplit` function splits a `data.frame` into parts, either via a
training/test split, or a K-fold split.
```{r}
df <- data.frame(matrix(
  rnorm(100 * 4), 100, 4
))
```
```{r}
split1 <- dfSplit(df, method = 'traintest', train_frac = 0.8)
lapply(split1, dim)

split2 <- dfSplit(df, method = 'traintest', train_n = 70)
lapply(split2, dim)

split3 <- dfSplit(df, method = 'kfold', K = 4)
lapply(split3, dim)
```

### The `fullRanker` Function
The `fullRanker` function takes a matrix or `data.frame` and successively
removes columns until it is of full rank.
```{r}
X1 <- diag(4)
X2 <- cbind(X1, 1)
```
```{r}
fullRanker(X1)
fullRanker(X2)
```

### The `unpack` Function
The `unpack` function simply takes objects in a list, and unpacks them into
the global (or otherwise unspecified) environment.
```{r}
rm(list = ls())
ls()
mylist <- list(a = 1:5,
               b = letters[1:6],
               c = rnorm(7))
unpack(mylist)
ls()
```
